A few stickleback suffice to transport adaptive alleles to new lakes

### 2018-12-13 Stickleback_Paper.tex
peter edits until disco

\paragraph{Habitat and geography}
Throughout the simulation there are two habitat types -- Marine and Freshwater-- defined by their selective pressures.
We start the simulations with two populations of size 5,000 diploid individuals in each of the habitats, for a total of 10,000 diploid individuals.
The arrangement of these habitats, depicted in Figure \ref{fig:Geo},
roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units,
while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
each connected to the marine habitat at regularly spaced intervals
(positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait
with different optima in marine and freshwater habitats.
This situation roughly models the cumulative effect of the various phenotypes such as armor morphology, 
body size, craniofacial variation and opercle shape on which divergent selection is thought to act on in the two environments. 
Concretely, the optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
is determined by a Gaussian kernel with standard deviation 15, i.e.,
\begin{align*}
    f(x_\text{ind}; x_\text{opt})
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x_\text{ind}-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}

In short, the difference between an individual's trait value and the optimum largely determines that individual's fitness. 
Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

\begin{figure}
	\begin{center}
  		\includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
  		\caption{
            \textbf{Diagram of simulated populations:}
            a single, continuous, one-dimensional marine habitat (blue)
            is coupled to randomly mating ``lakes'' at discrete locations with dark green arrows representing migration patterns.
            After an initial period of 100K generations with 25 lakes,
            an additional 25 lakes are added (at the same set of locations) and populated with a copy of the \textit{marine} individuals
            to simulate the appearance of newly accessible freshwater habitats colonized by marine stickleback.
            The marine habitat, and each set of 25 lakes, each contain 5,000 individuals at all times.
			}
  		\label{fig:Geo}
	\end{center}
\end{figure}

\paragraph{Genetic architecture of the trait}
Each individual carries two linear chromosomes, each of size $10^8$ loci.
Mutations that can affect the trait under selection can occur at rate $10^{-7}$ per locus per generation
in ten ``effect regions'' of $10^5$ loci each,
spread evenly along the chromosome.
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value 
$x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
Subsequent mutations at the same locus replace the previous allele.

\paragraph{Population dynamics}
We use SLiM to simulate a Wright--Fisher population with non-overlapping generations 
and a fixed population size of 5,000 diploid individuals in each habitat.
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(all individuals are hermaphroditic),
and the contributing genomes are produced by Poisson recombination 
with an average of one crossover per chromosome per generation 
($10^{-7}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000,
we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
to do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, 
so that the mean fitnesses of all lakes are equal before selection happens.

As shown in Figure~\ref{fig:Geo} as lines and arrows, dispersal occurs both locally along the coastline in the marine habitat
as well as between the marine habitat and the lakes.
All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
The lake--ocean migration rate is denoted $m$, 
which we refer to as the rate of \emph{gene flow} between habitats.
Each new individual in each habitat 
has parents from the other habitat with probability $m$ 
(in which case we call it a ``migrant''),
and parents from the same habitat with probability $1-m$.
% how do we get nonmigrants?
The first parent of each nonmigrant individual in the freshwater habitat
is chosen from the freshwater habitat proportional to fitness, 
and a mate is chosen from the same lake as the first, also proportional to fitness. 
The offspring then lives in the same lake as the parents.
Parents for each nonmigrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, 
then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
and reflected to stay within the population.
% what about migrants?
Parents for each freshwater migrant are chosen in the same way
as for nonmigrant marine individuals,
and are assigned to the lake nearest to the position of the first marine parent.
Similarly, parents for each marine migrant are both chosen from the same lake as before,
and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

\paragraph{New freshwater populations} 
To study how newly appearing freshwater habitats adapt and select on the standing genetic variation in the marine, we introduce a new set of 25 lakes midway through the simulation at 100K generations.
These new lakes are populated with a copy of the marine individuals to emulate a freshwater lake being colonized by its neighboring marine population. 
In particular, after this point of introduction in the simulation there are:
two \emph{sets} of lakes, and one marine population, each with 5,000 individuals for a total of 15,000 individuals being tracked in the simulation. 
Since this introduction of lakes doubles the number of lake-to-marine immigrants, 
the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.

\paragraph{Descriptive statistics}\bill{Population genetic analyses? Instead of the more general 'Descriptive statistics'?}
To assess whether new lakes adapt using existing genetic diversity,
we define a \emph{freshwater allele}
to be an effect mutation that has frequency higher than $0.5$ in at least one of the original lakes,
while remaining lower than $0.5$ in the marine. 
This categorization is made for each generation using the allele frequencies from that generation,
and so changes with time.
Alleles common in the newly introduced lakes do not count
if they are not also common in the original lakes.
They are defined this way because the transportation hypothesis 
does not specify where or when an advantageous mutation arises,
but simply suggests that any sufficiently common freshwater adapted allele 
could participate in adaptation in new habitat \citep{schluter2009genetics}.

\emph{Time to adaptation} of the introduced population, denoted $T_\text{adapt}$, 
is defined to be the generation at which 
the difference between the average trait value in the original and the introduced freshwater populations 
is less than 0.5.

We describe overall genetic differentiation between the habitats using $F_{ST}$,
calculated on a per-locus basis.
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
in the freshwater and marine habitats, respectively,
and $\bar p = (p_f + p_m)/2$,
then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.

\paragraph{Recording genealogical history}
We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
to output the genealogical history of all individuals
at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation.
This allowed us to directly query the true origins of adaptive alleles.
In addition it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations
which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}).

The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome.
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
\begin{enumerate}
    \item \emph{a ``De novo" allele:}
        deriving from a new mutation that occurred in a new lake.
    \item \emph{a ``Migrant" allele:} 
        deriving from a migrant that was not in the initial generation that colonized the lake
    \item \emph{a ``Captured" allele:}
        present in the individual that initially colonized the new lake, 
        and both common (above 50\%) in the original lakes,
        and uncommon (below 50\%) in the ocean.
    \item \emph{a ``Marine" allele:} 
        present in the individual that initially colonized the new lake, 
        and not a captured allele.
\end{enumerate}

We then looked at the ratio of alleles from each origin to the total number of alleles being used as the basis for local adaptation in the new lakes.
This essentially gives a measure on whether selection in the new environments made use of
(1) new mutation, 
(2) post-colonization migration, 
(3) standing variation at migration--selection balance, and
(4) standing variation at mutation--selection balance.
In other words, we were able to quantify the contributions of these possible origins 
of adaptive alleles to understand where the genomic basis of adaptation had derived from.

We were also able to use the tree sequence to get information about the individual ``effect regions''
in all initial genomes of the new lakes at time of introduction.
From this we determined the ability for the new lakes to select upon the standing variation in the marine.
Given the probability that effect region fixes at a given trait value, $2 * x / 45$, where $x$ is the effect size of a haplotype,
we calculated the expected total effect sizes of the haplotypes that fix.
Concretely, it is the sum across the 10 effect regions of
$$\sum_{i=1}^N \prod_{j < i} (1 - p_j) p_i x_i$$
where $N$ is the number of genomes per lake, $p_i$ is the probability of fixation of the 
$i^{th}$ haplotype, $x_{i}$ is the effect size of the $i^{th}$ haplotype, and the haplotypes are sorted in decreasing order by $x_{i}$.
These numbers were computed assuming additivity of mutations, meaning the numbers produced are a slight overestimate.

\section*{Results}

To observe the impact of gene flow on selection in the new lakes, we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
Since each lake contains 200 individuals, we refer to these migration rates in terms of \emph{dispersal} ($M$) \jgg{($D?$)} ranging from $0.01$ to $100$ migrants per lake, per generation.
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
At very low rates of gene flow, each new lake's population adapted completely separately from \emph{de novo} mutation, which took a very long time ($\approx 20,000$ generations).
At very high rates of gene flow, local adaptation was constrained by the large influx of maladaptive alleles.
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
Surprisingly, only a few migrants per generation from the lakes to the marine were needed
to see the effects of rapid local adaptation accelerated by selection on standing genetic variants in the new lakes -- despite being deleterious in the intervening marine habitat.

\subsection*{Local Adaptation: differentiation with gene flow}

@@

\begin{figure}
	\begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
  		\caption{ 
        		Mean individual trait values in the marine habitat (blue line),
        		the original lakes (yellow lines; average in orange),
        		and the new lakes (light green lines; average in dark green),
        		across the course of two simulations, with migration rates of
        		\textbf{(top)} $m=5 \times 10^{-5}$ and
        		\textbf{(bottom)} $m=5 \times 10^{-4}$ per generation per individual
        		(i.e., 0.01 and 0.1 migrants per lake per generation, respectively).
        		Optimal trait values in the two habitats are at $\pm 10$.
                	Analogous plots for other migration rates
                	are shown in Figure \ref{fig:phenotype_time_supp}.
		}
  		\label{fig:phenotype_time}
	\end{center}
\end{figure}

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred at migration rates of $m \le 10$.
% with the exception of the highest gene flow at which half of each population was composed of migrants.
%At lower migration rates (10 migrants per generation and below), populations adapted to local conditions.
Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}
show freshwater and marine populations diverging until the trait means were close to the optimal values in each habitat. 
The establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value.
%which move the trait by an amount of order 1 every few hundred generations.
%Trait variation within each population was small compared to the difference between populations.
At the lowest rate of gene flow, $m = 0.01$, differences at $\approx 20$ commonly polymorphic sites 
($\approx 10$ that shift the trait in each direction)
were responsible for most of the adaptive differences between freshwater and marine habitats.
In Figure \ref{fig:MPFAI}A, we can see that more gene flow simplified the genetic basis. %does this belong here?
As expected, increasing migration rate decreased differentiation between habitats. 
As seen in Figure~\ref{fig:Fst}, 
$F_{ST}$ between marine and freshwater habitats 
at neutral sites steadily declines as migration increases.

@@

\paragraph*{Speed of adaptation}
Adaptation occurred much more quickly
at higher migration rates,
both in the old and new sets of lakes.
We measured this ``time to adaptation''
as the number of generations until 
average trait values in old and new lakes were within 0.5 of each other,
shown in Figure \ref{fig:TimeToAdaptation} for different rates of gene flow.
Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$,
while at $m = 1$ migrant per, lake per generation,
new lakes managed to adapt in just under 60 generations. 
%This dramatic decrease in time to local adaptation is suggestive of just how quickly
%selection can act on standing variation. 
%It also suggests a rough threshold of gene flow required 
%to efficiently select upon the alleles present in the marine individuals.

@@

\subsection*{Sharing of freshwater adapted alleles}

At low migration rates, the \emph{initial} period of adaptation
takes roughly 25 times longer for lakes than it does for the ocean.
This difference occurs because at low migration rates,
all subpopulation must wait for novel mutations to arise in order to adapt.
Because the marine is one continuous population with 25 times more individuals than any one lake, 
there is a much larger influx of new mutations to be selected upon. 
At higher migration rates, greater mixing
allows the initial lakes to share alleles instead of developing their own genetic basis of adaptation. 
As a first indication of this, Figure \ref{fig:Fst} 
shows that $F_{ST}$ between the ``original'' and ``introduced'' sets of lakes 
at effect mutations decreased with migration rate.
%We are interested in this shared genetic basis of adaptation because it is commonly thought that this is
%a requirement for the efficient ``transportation" of the freshwater haplotype in to new lakes \citet{schluter2009genetics}

To investigate in more depth how locally adaptive alleles found in the original lakes
are shared between lakes, as well as how they spread to the new lakes,
we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
To be considered in this category, a mutation must participate in the genetic basis of any freshwater adaptation.
Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lakes and below 50\% in the marine habitat.
Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
At $m = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere:
the new lakes acquire new adaptive alleles and therefor have little to none of these adaptive alleles.
At $m = 0.1$, we again observe the original lakes adapting nearly independently from each other.
However, in Figure \ref{fig:MPFAI}A we can see the average marine individual is shown to hold $\approx 2$ pre-existing freshwater adapted alleles. 
This being the case, we can see new lakes adapt using a subset of this large repertoire of standing variation deriving from a multitude of independently derived haplotypes.
As migration rate increases past this, %we can see migration selection balance keeps the average capacity for marine individuals to carry freshwater alleles at about 2.
we can see in Figure \ref{fig:MPFAI}C that the total number of pre-existing freshwater adapted alleles declines.
%while the number of alleles which make up the genetic basis of adaptation in freshwater individuals remains relatively similar.
%Although the genetic basis seems to simplify slightly due to more , this 
This suggests that the alleles move between populations by migration 
before selection acts on new mutation for neighboring lakes.
Interestingly, the frequency of these alleles in the ocean
stays relatively constant across the reasonable rates of gene flow.

Figure~\ref{fig:MPFAI}B shows the distribution, through time, of 
the mean percentage of currently-defined freshwater adapted alleles 
that each genome in each of the populations carries.
If all individuals across lakes carried the same set of alleles determining their trait value, 
this would be 100\%. 
At the lowest migration rate $m = 0.01$,
each genome in the original lakes have almost exactly~$1/25^{th}$ 
of the total number of pre-existing freshwater adapted alleles -- 
this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above,
the average individual across the new lakes has nearly the same amount of 
pre-existing freshwater adapted alleles as individuals across the old lakes.
As expected, the genetic basis of the freshwater phenotype 
seems to simplify as migration increases --
higher rates of migration allow
adaptive alleles of higher effect to travel more efficiently through the population,
even though they are deleterious in the ocean.

The numbers of Figure~\ref{fig:MPFAI} suggest that
the dramatic increase in speed of local adaptation we observed above occurs because
higher gene flow between populations allows sharing of freshwater alleles between populations.
%however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
%For selection to act on SGV on the scale we observe in nature, the gene flow 
%initial generation to be able to rebuild 
We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting
allele, in each individual, common in any of the new lakes at time of adaptation
as defined in the Methods.
Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo"} mutation.
As gene flow increases, a larger fraction of adaptive alleles 
derive from pre-existing variation in the marine population at the time of introduction.
In other words, 
greater mixing at higher migration rates allows lakes to share alleles
instead of developing their own genetic basis of adaptation.

@@

\subsection*{Realized genetic architecture}

Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
Do measures of local differentiation such as $F_{st}$ scans or Genome Wide Association Studies (GWAS) identify the causal loci? 
Figure \ref{fig:Fst} shows plots along the genome of per-locus $F_{ST}$ values between the marine habitat and all freshwater habitats pooled together.
Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci 
``Background'' levels of $F_{ST}$ increase as gene flow decreases,
swamping out this signal until the regions under selection are indistinguishable. 
This is likely the contribution of two separate forces of natural selection: 
first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
second, greater sharing of adaptive alleles providing a shared signal across populations.

>>>>

\paragraph{Habitat and geography}
Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures,
each with 5,000 diploid individuals.
The arrangement of these habitats, depicted in Figure~\ref{fig:Geo},
roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units,
while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
each connected to the marine habitat at regularly spaced intervals
(positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait
with different optima in marine and freshwater habitats.
This situation roughly models the cumulative effect of the various phenotypes 
thought to be under divergent selection between the habitats,
such as armor morphology, body size, craniofacial variation and opercle shape.
The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
is determined by a Gaussian kernel with standard deviation 15, i.e.,
\begin{align*}
    f(x_\text{ind}; x_\text{opt})
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x_\text{ind}-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}
In short, the difference between an individual's trait value and the optimum determines that individual's fitness. 
% Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

\begin{figure}
    \begin{center}
          \includegraphics[width=0.6\linewidth]{GeographyFigure.pdf}
          \caption{
            \textbf{Diagram of simulated populations:}
            a single, continuous, one-dimensional marine habitat (blue)
            is coupled to randomly mating ``lakes'' at discrete locations with arrows representing migration patterns.
            After an initial period of 100K generations with 25 lakes,
            an additional 25 lakes are added (at the same set of locations) and populated with marine individuals
            to simulate the appearance of newly accessible freshwater habitats colonized by marine stickleback.
            The marine habitat, and each set of 25 lakes, each contain 5,000 individuals at all times.
            }
          \label{fig:Geo}
    \end{center}
\end{figure}

\paragraph{Genetic architecture of the trait}
Each individual carries two linear chromosomes, each of size $10^8$ loci.
Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation
in ten regions of $10^5$ loci each,
spread evenly along the chromosome.
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value 
$x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
Subsequent mutations at the same locus replace the previous allele.
We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation, 
as described in \citet{kelleher2018efficient}.

\paragraph{Population dynamics}
We use SLiM to simulate a Wright--Fisher population with non-overlapping generations 
and a fixed population size of 5,000 diploid individuals in each habitat.
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(all individuals are hermaphroditic),
and the contributing genomes are produced by Poisson recombination 
with an average of one crossover per chromosome per generation 
($10^{-7}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000,
and the Wright--Fisher model assumes unrealistic global population regulation,
we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
to do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, 
so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat
as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$.
All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat 
has parents from the other habitat with probability $m$ 
(in which case we call it a ``migrant''),
and parents from the same habitat with probability $1-m$.
% how do we get nonmigrants?
The first parent of each nonmigrant individual in the freshwater habitat
is chosen from the freshwater habitat proportional to fitness, 
and a mate is chosen from the same lake as the first, also proportional to fitness. 
The offspring then lives in the same lake as the parents.
Parents for each nonmigrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, 
then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
and reflected to stay within the habitat range.
% what about migrants?
Parents for each freshwater migrant are chosen in the same way
as for nonmigrant marine individuals,
and are assigned to the lake nearest to the position of the first marine parent.
Similarly, parents for each marine migrant are both chosen from the same lake as before,
and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

\paragraph{New lakes} 
To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats,
we introduce a new set of 25 lakes after 100,000 generations.
These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by
oceanic stickleback.
This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals.
Since this introduction of lakes doubles the number of lake-to-marine immigrants, 
the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.

\paragraph{Recording genealogical history}
We used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
to output the genealogical history of all individuals
at the time of introduction of new lakes, at the time of adaptation, and at the end of the simulation.
This allowed us to directly query the true origins of adaptive alleles.
In addition, it allowed for much larger simulations by avoiding the computationally expensive task of simulating neutral mutations
which were retroactively added to the gene trees at a rate of $10^{-7}$ per locus per generation, as described in \citet{kelleher2018efficient}.

The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome.
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
\begin{enumerate}
    \item \emph{a ``De novo" allele:}
        deriving from a new mutation that occurred in a new lake.
    \item \emph{a ``Migrant" allele:} 
        deriving from a migrant not in the initial generation that colonized the lake
    \item \emph{a ``Captured" allele:}
        present in initial colonists of the new lake, 
        and both common (above 50\%) in the original lakes,
        and uncommon (below 50\%) in the ocean.
    \item \emph{a ``Marine" allele:} 
        present in initial colonists of the new lake, 
        but not a ``captured'' allele.
\end{enumerate}
The proportion of trait-affecting alleles in new lakes that fall in these categories
measures the degree to which selection in the new environments made use of
(1) new mutation, 
(2) post-colonization migration, 
(3) standing variation at migration--selection balance, and
(4) standing variation at mutation--selection balance.

We also used neutral mutations to calculate $F_{ST}$,
calculated on a per-locus basis and then averaged in windows 
as a measure of between-population relative differentiation.
Concretely, if $p_f$ and $p_m$ are the frequencies of a given mutant allele
in the freshwater and marine habitats, respectively,
and $\bar p = (p_f + p_m)/2$,
then we compute $F_{ST}$ for that mutation as $1 - p_f p_m / (\bar p (1-\bar p))$.

%%%%%%%%%%%%%%%%%%
\section*{Results}

To observe the impact of gene flow on selection in the new lakes, 
we varied the ocean--lake migration rate, $m$, across separate simulations from $5 \times 10^{-5}$ to $5 \times 10^{-1}$.
Below, we often refer to these migration rates in terms of the number of migrants per lake, per generation, which we denote $M$.
Since each lake contains 200 individuals, $M=200m$.
Many aspects of adaptation changed substantially across this range, including the speed of adaptation, 
degree of sharing of adaptive alleles between lakes, and the population genetic signals left behind.
At very low rates of gene flow, each new lake's population adapted almost completely independently throught \emph{de novo} mutation, 
which took a very long time ($\approx 20,000$ generations).
At very high rates of gene flow, local adaptation was constrained by the large influx of locally maladaptive alleles.
Between these two extremes, genetic variation that allowed adaptation to freshwater habitats could move relatively easily between lakes.
Perhaps surprisingly, only a few migrants per generation from the lakes to the ocean were needed
to maintain sufficient genetic variation in the ocean to accelerate adaptation in new lakes.

%%%%%%%%%%%%
\subsection*{Local Adaptation: differentiation with gene flow}

@@

\begin{figure}
    \begin{center}
        \includegraphics{Final_Plots/Pheno_Time.pdf}
          \caption{ 
                Mean individual trait values in the marine habitat (blue line),
                the original lakes (yellow lines; average in orange),
                and the new lakes (light green lines; average in dark green),
                across the course of two simulations, with migration rates of
                \textbf{(top)} $M=0.01$ and
                \textbf{(bottom)} $M=0.1$ 
                migrants per lake per generation, respectively.
                Optimal trait values in the two habitats are at $\pm 10$.
                    Analogous plots for other migration rates
                    are shown in Figure \ref{fig:phenotype_time_supp}.
        }
          \label{fig:phenotype_time}
    \end{center}
\end{figure}

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations
with the exception of the highest gene flow at which half of each population was composed of migrants.
Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}
show how mean trait values in freshwater and marine populations diverged over time,
until the trait means were close to the optimal values in each habitat. 
Establishment of new alleles in the lakes is visible 
in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value
which move the trait by an amount of order 1 every few hundred generations.
%Trait variation within each population was small compared to the difference between populations.
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites 
(about 10 that shift the trait in each direction)
were responsible for most of the adaptive differences between freshwater and marine habitats.
As expected, increasing migration rate decreased differentiation between habitats:
as seen in Figure~\ref{fig:Fst}, 
$F_{ST}$ between marine and freshwater habitats 
at neutral sites steadily declines as migration increases.

@@

Adaptation occurred much more quickly
at higher migration rates,
both in the old and new sets of lakes.
We measured this ``time to adaptation''
as the number of generations until 
average trait values in old and new lakes were within 0.5 of each other,
shown in Figure~\ref{fig:TimeToAdaptation} for different rates of gene flow.
Adaptation of new lakes took over 18,000 generations at the lowest rate of $m = 0.01$,
while at $m = 1$ migrant per, lake per generation,
new lakes managed to adapt in just under 60 generations.

@@

%%%%%%%%%%%%
\subsection*{Sharing of freshwater adapted alleles}

At low migration rates, the \emph{initial} period of adaptation
takes roughly 25 times longer for lakes than it does for the ocean.
This difference occurs because at low migration rates,
each lake must wait for its own novel mutations to arise in order to adapt.
Because the marine habitat is continuous, with 25 times more individuals than any one lake, 
there is a much larger influx of new mutations to be selected upon. 
At higher migration rates, greater mixing
allows the initial lakes to share alleles instead of developing their own genetic basis for adaptation.

To investigate in more depth how locally adaptive alleles found in the original lakes
are shared between lakes, as well as how they spread to the new lakes,
we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation.
To be considered in this category, 
an allele must participate in the genetic basis of local adaptation for at least one of the original lakes.
Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat.
Figure \ref{fig:MPFAI}A shows the distribution of the number of these alleles across generations.
At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere:
new lakes independently acquire new adaptive alleles rather than pre-existing ones.
At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other,
but now the new lakes adapt using pre-existing alleles present in the original set of lakes.
Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles,
standing variation which was nearly absent at $M=0.01$.
As migration rate increases past this, 
the total number of pre-existing freshwater adapted alleles declines (Figure \ref{fig:MPFAI}C).
Interestingly, the frequency of these alleles in the ocean
stays relatively constant across the reasonable rates of gene flow.

Figure~\ref{fig:MPFAI}B shows the distribution, through time, of 
the mean \emph{percentage} of currently-defined freshwater adapted alleles 
that each genome in each of the populations carries.
If all individuals across lakes carried the same set of alleles determining their trait value, 
this would be 100\%. 
At the lowest migration rate $m = 0.01$,
each genome in the original lakes have almost exactly~$1/25^{th}$ 
of the total number of pre-existing freshwater adapted alleles -- 
this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes.
Figure \ref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above,
the average individual across the new lakes has nearly the same amount of 
pre-existing freshwater adapted alleles as individuals across the old lakes.
As expected, the genetic basis of the freshwater phenotype 
seems to simplify as migration increases --
higher rates of migration allow
adaptive alleles of larger effect to travel more efficiently through the population,
even though they are deleterious in the ocean.

The numbers in Figure~\ref{fig:MPFAI} suggest that
the dramatic increase in speed of local adaptation we observed above occurs because
higher gene flow between populations allows sharing of freshwater alleles between populations.
%however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
%For selection to act on SGV on the scale we observe in nature, the gene flow 
%initial generation to be able to rebuild 
We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele
in each individual in the new lakes, as defined in the Methods.
Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation.
As gene flow increases, a larger fraction of adaptive alleles 
derive from pre-existing variation in the marine population at the time of introduction.
In other words, 
greater mixing at higher migration rates allows lakes to share alleles
instead of developing their own genetic basis of adaptation.

@@

%%%%%%%%%%%%
\subsection*{Scans for selection}

Here, we take a closer look at the genomic architecture of local adaptation between the two habitats. 
Can measures of local differentiation such as $F_{ST}$ be used to identify the causal loci? 
Figure \ref{fig:Fst} shows plots along the genome of average per-locus $F_{ST}$ values 
in 500bp windows between the marine habitat and all freshwater habitats pooled together.
\plr{500bp windows, right?}
Higher rates of dispersal showed more distinct $F_{ST}$ peaks over polymorphic loci,
while ``background'' levels of $F_{ST}$ increase as gene flow decreases,
swamping out this signal until the regions under selection are indistinguishable. 
This is likely the contribution of two separate forces of natural selection: 
first, stronger genetic drift with less migration leads to higher background $F_{ST}$, and 
second, greater sharing of adaptive alleles providing a shared signal across populations.
